Toward Adaptive GPU Resource Management for Embedded Real-Time Systems

### 2012-03-09 draft/draft.tex
The figure is added, and the tex file is also modified.

\author {
Junsung Kim and Ragunathan (Raj) Rajkumar\\
\\
Department of Electrical and Computer Engineering\\
Carnegie Mellon University
\and
Shinpei Kato and Scott Brandt\\
\\
Department of Computer Science\\
University of California, Santa Cruz
}

@@

\section{Introduction}
\label{sec:introduction}
Graphics processing units (GPUs) are widely used in a range of applications from supercomputers to mobile computers. For example, three of the top five supercomputers on the TOP500 list~\cite{TOP500} utilize GPUs to accelerate computations, and recent tablets such as ASUS Eee Pad Transformer Prime leverage 12 GPU cores of NVIDIA’s well-known SoC, Tegra 3 to boost the performance while saving power~\cite{Tegra3}. This trend will only continue and in fact is expected to accelerate across multiple application domains.

Modern automobiles have several tens of processing units, and advances in safe-driving features such as adaptive cruise control, stop-and-go cruise control, lane keeping and assisted lane change require even higher performance computing. As vehicles become semi-autonomous, drive autonomously on demand and eventually become fully autonomous, a multitude of computer vision, sensor fusion, signal processing and graphics problems must be solved in real time. Since solutions to those problems can leverage benefits of parallel processing techniques, GPUs can be used to accelerate computations. As an initial step, NVIDIA GPUs will be used for infotainment systems in future BMW vehicles~\cite{NVIDIA_BMW}.

Recent trends in self-driving cars~\cite{Urmson08, Markoff10, Kelly12} are motivated by various benefits of automated driving, but most importantly enabling automatic safety features. Those features are realized by the smart vehicle planning and the intelligent processing of data from various types of sensors on self-driving cars such as LIDAR (LIght Detection And Ranging), radar, camera, ultrasonic sensor, and so forth. One common characteristic of both the smart vehicle planning and the intelligent data processing is that their performances can be greatly accelerated using GPU, hence parallelism. For example, the more a self-driving car plans its potential paths to follow {\it a priori}, the better the quality of self-driving is. Since the self-driving car should follow only the best path, calculation of each potential path can happen parallel. A CMU team has already started using GPUs for autonomous driving computations. They have increased the performance of motion planning for autonomous driving up to 40 times using an NVIDIA GPU GTX 260 having 192 logical cores~\cite{McNaughton11}.

The perception system which intelligently processes the data from various sensors can also obtain benefits using GPU. The perception system of a self-driving car should be able to detect, classify and track the obstacles around itself. Various types of sensors will generate huge amount of information to understand the surroundings. In order to process the data, a parallel processing technique is of vital importance. For example, a new self-driving car at CMU manages 1536 objects from LIDAR sensors before they are fused with other types of sensor data. There has been on-going research using GPU to build a perception system~\cite{Ferreira11}, and their GPU implementation yielded 30,000-times-faster performance compared to the case of using only CPU.

@@

In the following, we present two frameworks that could achieve our idea.
Although they are still at an idea stage, we plan to implement a real
system as a proof-of-concept, leveraging open-source
software~\cite{Kato_OSPERT11}.

>>>>

\author {
Junsung Kim and Ragunathan (Raj) Rajkumar\\
\\
Department of Electrical and Computer Engineering\\
Carnegie Mellon University
\and
Shinpei Kato\\
\\
Department of Computer Science\\
University of California, Santa Cruz
}

@@

\section{Introduction}
\label{sec:introduction}

Graphics processing units (GPUs) are getting more and more commonplace
in many application domains widely ranging from high-performance
computing to embedded mobile computing.
For example, three of the top five supercomputers on the TOP500
list~\cite{TOP500}, announced as of March 2012, use GPUs to accelerate
computations, while recent tablets, such as ASUS Eee Pad Transformer
Prime, also leverage embedded GPUs, \textit{a.k.a.}, Tegra
3~\cite{Tegra3}, to enhance performance under power constraints.
This trend will only continue and in fact is expected to appear across
more and more application domains.

One notable application domain of GPUs is automotive engineering.
Modern automobiles employ several tens of processing units.
Further advances in safe-driving features, such as adaptive
cruise control, stop-and-go cruise control, lane keeping, and assisted
lane change, would require even higher computing capabilities.
As vehicles are becoming autonomous, a multitude of computer vision,
sensor fusion, signal processing, and graphics systems must operate and
communicate in real-time.
Given their highly data-parallel and compute-intensive workloads, 
we believe that parallel computing is a useful solution.
As technology reads today, the GPU is the most well-suited platform.
In fact, NVIDIA GPUs will be used for infotainment systems platforms in
future product lines of BMW vehicles~\cite{NVIDIA_BMW}.

Recent trends in autonomous-driving vehicles~\cite{Kelly12, Markoff10,
Urmson08} are motivated by various benefits of automated functions,
but, most importantly, enabling automatic safety features. 
These features are often brought by smart planning and intelligent
processing of data obtained from many sensors equiped in the vehicle,
including LIDAR (LIght Detection And Ranging), radar, camera, and
ultrasonic sensors. 
One common characteristic in these types of processing is that GPUs can
accelerate their processing speeds significantly.
For example, autonomous driving should follow only the best path among
many potential paths whose calculations can happen in parallel.
Calculating as many potential paths to follow as possible
{\it a priori}, we will get better quality of driving.
As a matter of fact, CMU's autonomous vehicle team showed that their
motion planning algorithm was sped up by 40 times~\cite{McNaughton11},
using an NVIDIA GTX 260 GPU that integrates 192 compute cores on a
chip.

In addition to motion planning, a perception algorithm can benefit
from the GPU as well in sensor data processing.
The perception system of a self-driving car should
be able to detect, classify and track the obstacles around
itself. Various types of sensors will generate huge amount of
information to understand the surroundings. In order to process the
data, a parallel processing technique is of vital importance. For
example, a new self-driving car at CMU manages 1536 objects from LIDAR
sensors before they are fused with other types of sensor data. There has
been on-going research using GPU to build a perception
system~\cite{Ferreira11}, and their GPU implementation yielded
30,000-times-faster performance compared to the case of using only CPU.

@@

\begin{figure}[t]
\includegraphics[width=3.0in]{architecture.eps}
\caption{The proposed conceptual architecture for GPU resource management.} 
\label{fig_arch}
\end{figure}

In the following, we present two frameworks that could achieve our idea.
Although they are still at an idea stage, we plan to implement a real
system as a proof-of-concept, leveraging open-source
software~\cite{Kato_OSPERT11}. The proposed architecture is also 
illustrated in Figure~\ref{fig_arch}.
